Notes

### 2022-09-04 chapters/appendix-code.tex
Finish first complete revision

When possible, the code was implemented in a generic way that would allow for easy extensions or derivations, and also with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries. Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them. The code for the experiments was executed in a machine with 4 cores, and a random seed with value 2022 was set for reproducibility. We provide a script file \texttt{launch.sh} that illustrates a typical execution. Lastly, there are also Jupyter notebooks that demonstrate the use of our methods in a more visual way. Inside these notebooks there is a step-by-step guide on how one might execute our algorithms, accompanied by many graphical representations, and offering the possibility of changing multiple parameters to experiment with the code. In addition, there is also a notebook that can be used to generate all the tables and figures of this document pertaining to the experimental results.

>>>>

When possible, the code was implemented in a generic way that would allow for easy extensions or derivations. It was also developed with efficiency in mind, so many functions and methods exploit the vectorization capabilities of the \textit{numpy} and \textit{scipy} libraries. Moreover, we followed closely the style of the \textit{scikit-learn} and \textit{scikit-fda} libraries, so our methods are fully compatible and could be integrated (after some minor tweaking) with both of them. The code for the experiments was executed in a machine with 4 cores, and a random seed with value 2022 was set for reproducibility. We provide a script file \texttt{launch.sh} that illustrates a typical execution. Lastly, there are also Jupyter notebooks that demonstrate the use of our methods in a more visual way. Inside these notebooks there is a step-by-step guide on how one might execute our algorithms, accompanied by many graphical representations, and offering the possibility of changing multiple parameters to experiment with the code. In addition, there is also a notebook that can be used to generate all the tables and figures of this document pertaining to the experimental results.

### 2022-09-04 chapters/background-related-work.tex
Finish first complete revision

FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}. Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint. Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}. At the same time, much effort has also been put into the opposite task: generalizing different concepts to the infinite-dimensional framework. Examples of this endeavor include a functional Mahalanobis distance \citep{berrendero2020mahalanobis} or a functional partial least squares algorithm \citep{delaigle2012methodology}. Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}. These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep[e.g.][]{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems. Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.

@@

 The most common scalar-on-function linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^p\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\). A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}). This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study (see \citet{cardot2011functional} for more insights).

Technical details aside, the inference on \(\beta\) is also hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
Although it can be tempting to simply discretize the observed values on a grid and proceed with standard multiple linear regression, this would result in an under-determined model in which the estimated parameters lose meaning (essentially because the number of available parameters is infinite, but the number of equations is finite). Thus, some regularization or dimensionality reduction techniques are needed for parameter estimation; see \citet{reiss2017methods} for a summary of several widespread methods.

A common inference strategy is to expand both \(X\) and \(\beta\) on a certain data-driven orthonormal basis of \(L^2[0,1]\), say \(\{\phi_j\}\), up to a certain value \(p\in\N\):
\[
  X(t)=\sum_{j=1}^p Z_j \phi_j(t), \quad \beta(t) = \sum_{j=1}^p \beta_j \phi_j(t).
\]
In this way, model~\eqref{eq:l2-linear-model} simplifies to
\[
Y = \alpha_0 + \sum_{j=1}^p \beta_j Z_j,
\]
which can then be solved in the usual manner. For example, when the chosen basis is composed of eigenfunctions of the covariance operator \(\K\), this method is known as Functional Principal Component Regression (FPCR). Another class of methods focus on selecting an \textit{a priori} basis for \(\beta\) (e.g. a spline basis) and solving a penalized least squares problem. In particular, for a truncated basis representation of the form \(\beta(t)=\phi(t)'b\), with \(\phi(t)=(\phi_1(t), \dots, \phi_p(t))'\) and \(b\in\R^p\), a typical minimization problem would be
\[
\argmin_{(\alpha_0, b)\in\R\times\R^p} \ \left\{ \sum_{i=1}^n \left( y_i - \alpha_0 - \int_0^1 x_i(t)[\phi(t)'b]\, dt\right)^2 + \lambda \Omega(b) \right\},
\]
where \(\Omega\) is some penalty function and \(\lambda>0\) is a regularization parameter.

@@

The term Bayesian inference usually refers to a wide class of methods that to some extent employ Bayes' theorem to update the initial probability assigned to a hypothesis when new information is available. It can be seen as a general tool for modeling situations or problems that involve uncertainty; some examples include Bayesian hierarchical modeling, Bayesian regression or even Bayesian neural networks \citep[see e.g.][]{murphy2012machine, bishop2006pattern}. Specifically, in this work we will be interested in performing parameter estimation in a Bayesian framework, so that pre-existing information and beliefs about the parameters can be incorporated into the model.

@@

In this section we present a brief exposition of some basic concepts regarding reproducing kernel Hilbert spaces. Although there are quite a few ways of introducing these spaces, we adopt a probabilistic point of view that will set the stage for the development of the subsequent theory. For a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}.

We start by defining the concept of kernel functions, which are the foundation of these spaces. Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity.

@@

Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and without loss of generality assume \(\E[X(t)]=0\) for all \(t\in[0,1]\). Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\). To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional vector space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
\begin{equation}\label{eq:h0}
\Hcal_0(K) = \left\{ f: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
\end{equation}
Note that, as subsets, \(\Hcal_0(K)\subset L^2[0,1]\). However, this new space can be endowed with an inner product different from the one induced by \(L^2[0,1]\), namely
\[
\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j),
\]
for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). We show below that it is well defined, but before we note that functions in this space satisfy the so-called \textit{reproducing property}: \(f(t) = \dotprod{K(t, \cdot)}{f}_K\) for all \(t \in [0, 1]\). In particular, \(K(t, s)=\dotprod{K(t, \cdot)}{K(s, \cdot)}_K\), which can be understood as saying that the kernel \textit{reproduces} itself, hence the term ``reproducing kernel''.

@@

\begin{example}[Non-RKHS Hilbert space]
  As we pointed out before, the space of square integrable functions \(L^2(\R)\) \textit{is not} a RKHS. There are many additional ways of seeing this; for example, in this space norm convergence does not imply pointwise convergence. Another possibility is to observe that the would-be reproducing kernel must be the Dirac delta function:
  \[
    f(s) = \int_{-\infty}^\infty f(t)\delta(t, s)\, dt \quad \forall s \in \R.
  \]
  However, \(\delta(t, \cdot) \notin L^2(\R)\).
\end{example}

@@

Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning. For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property, where a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS. Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
\[
K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\Hcal}.
\]
The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with an simple expression but a complex feature space (e.g. infinite-dimensional), such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).

@@

\begin{theorem}[Generalized representer theorem]
  Consider a RKHS \(\Hcal(K)\) of functions defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\). Then, any minimizer \(f^\ast\) of the regularized empirical risk
  \[
    R(\{(x_i, y_i, f(x_i))\}_{i=1}^n) + \Omega(\|f\|_K)
  \]
  is of the form \(f^\ast(\cdot) = \sum_{i=1}^n a_i K(x_i, \cdot)\), where \(a_i\in\R\) for all \(i=1,\dots,n\).
\end{theorem}

@@

The Metropolis-Hastings algorithm \citep{metropolis1953equation} is arguably the most well-known algorithm in this context. As a matter of fact, it is actually more of a general framework from which many other methods are derived. In its original formulation, it relies on a symmetric \textit{proposal} distribution \(g(x'|x_t)\) that generates a proposal for the next state given the current one (e.g. a Gaussian distribution centered on \(x_t\)). Approximate samples from the intended distribution can then be generated through the following iterative acceptance-based procedure:
\begin{enumerate}[1.]
  \item Generate a proposal \(x' \sim g(x'|x_t)\).
  \item Accept the proposal with probability \(\alpha=\min\{1, f(x')/f(x_t)\}\). If accepted, set \(x_{t+1}=x'\); otherwise set \(x_{t+1}=x_t\).
\end{enumerate}

@@

Figure~\ref{fig:mh} shows a schematic representation of the process. Note that \(f(x')/f(x_t)=\pi(x')/\pi(x_t)\) is a measure of how more likely is \(x'\) to be a sample from \(\pi\) than \(x_t\). This specific acceptance rate is also chosen so that the detailed balance equations of the chain are satisfied. An immediate generalization consists on choosing a general (possibly non-symmetric) jump distribution \(g\), in which case the acceptance probability is rewritten as
\[
\alpha = \min\left\{1, \frac{f(x')g(x_t \mid x')}{f(x_t)g(x'\mid x_t)}\right\}.
\]

Although this algorithm performs well on a wide range of problems, it can be difficult to find a suitable jump distribution when the number of dimensions is high. If we know the conditional distribution of each variable, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension. Specifically, in each step we get unidimensional samples \textit{in turn} (e.g. via rejection sampling) of each variable conditional on the rest of them, using always the more up-to-date values available:

>>>>

FDA is undoubtedly an active area of research, which finds applications in a wide variety of fields, such as biomedicine, finance, meteorology or chemistry \citep[see for example][]{ullah2013applications}. Accordingly, there are many recent contributions on how to tackle functional data problems, both from a theoretical and practical standpoint. Chief among them is the approach of reducing the problem to a finite-dimensional one, for example using a truncated basis expansion or spline interpolation methods \citep[e.g.][]{muller2005generalized, aguilera2013comparative}. At the same time, much effort has also been put into the task of building a sound theoretical basis for FDA, generalizing different concepts to the infinite-dimensional framework. Examples of this endeavor include the definition of centraility measures and depth-based notions for functional data \citep[e.g.][]{fraiman2001trimmed, cuevas2007robust, lopez2009concept}, an ANOVA test for functional data \citep{cuevas2004anova}, a purely functional partial least squares algorithm \citep{delaigle2012methodology}, a functional Mahalanobis distance \citep{berrendero2020mahalanobis}, or an extension of Fisher's discriminant analysis for function-valued random elements \citep[e.g.][]{james2001functional, shin2008extension}, among many others. Interestingly enough, these last two are examples of situations in which a RKHS-based approach provides useful insights.

Another technique found in the related literature is the use of Gaussian processes to model the functional behavior of the data \citep[see for instance][]{shi2011gaussian}. These ideas extend the theory of Gaussian process regression in classical finite-dimensional settings \citep[e.g.][]{rasmussen2004gaussian}, providing an alternative Bayesian approach to functional inference and prediction problems. Additional non-parametric methods for functional prediction and classification were notably explored in \citet{ferraty2006nonparametric}.

@@

 The most common scalar-on-function linear regression model is the classical \(L^2\)-model, widely popularized since the first edition (1997) of the monograph by~\citet{ramsay2005functional}. It can be seen as a generalization of the usual finite-dimensional model, replacing the scalar product in \(\R^p\) for that of the functional space \(L^2[0,1]\):
\begin{equation}\label{eq:l2-linear-model}
Y = \alpha_0 + \dotprod{X}{\beta} + \epsilon = \alpha_0 + \int_0^1 X(t)\beta(t)\, dt + \epsilon,
\end{equation}
where \(\alpha_0\in \R\), \(\epsilon\) is a random error term independent from \(X\) with \(\E [\epsilon]=0\), and the functional slope parameter \(\beta=\beta(\cdot)\) is assumed to be a member of the infinite-dimensional space \(L^2[0, 1]\). A careful rearrangement of~\eqref{eq:l2-linear-model} shows that the model is equivalently expressed as \(\Delta=\K\beta\), where \(\Delta(\cdot)\) is the cross-covariance function of \(X\) and \(Y\), and \(\K\) is the covariance operator of the process \(X\) (to be defined in Section~\ref{sec:rkhs}). This expression is a continuous analog of the normal equations that arise in finite-dimensional regression settings, and since the operator \(\K\) is non-invertible in general, questions of existence and unicity of \(\beta\) need further study \citep[see][]{cardot2011functional}.

Technical details aside, the inference on \(\beta\) is also hampered by the fact that \(L^2[0,1]\) is an extremely wide space that contains many non-smooth or ill-behaved functions, so that any estimation procedure involving optimization on it would typically be hard.
Although it can be tempting to simply discretize the observed values on a grid and proceed with standard multiple linear regression, this would result in an under-determined model in which the estimated parameters lose meaning. This happens essentially because the number of available parameters is infinite, but the number of equations is finite \citep[see][Sec.~15.2]{ramsay2005functional}. Thus, some regularization or dimensionality reduction techniques are needed for parameter estimation; see \citet{reiss2017methods} for a summary of several widespread methods.

A common inference strategy is to expand both \(X\) and \(\beta\) on a certain data-driven orthonormal basis of \(L^2[0,1]\), say \(\{\phi_j\}\), up to a certain value \(p\in\N\):

\[
  X(t)=\sum_{j=1}^p Z_j \phi_j(t), \quad \beta(t) = \sum_{j=1}^p \beta_j \phi_j(t).
\]
In this way, model~\eqref{eq:l2-linear-model} simplifies to
\[
Y = \alpha_0 + \sum_{j=1}^p \beta_j Z_j,
\]
which can then be solved in the usual manner. For example, when the chosen basis is composed of eigenfunctions of the covariance operator \(\K\), this method is known as Functional Principal Component Regression (FPCR). Another class of methods focus on selecting an \textit{a priori} basis for \(\beta\) (e.g. a spline basis) and solving a penalized least squares problem. In particular, for a truncated basis representation of the form \(\beta(t)=\phi(t)'b\), with \(\phi(t)=(\phi_1(t), \dots, \phi_p(t))'\) and \(b\in\R^p\), a typical minimization problem would be
\[
\argmin_{(\alpha_0, b)\in\R\times\R^p} \ \left\{ \sum_{i=1}^n \left( y_i - \alpha_0 - \int_0^1 x_i(t)[\phi(t)'b]\, dt\right)^2 + \lambda \Omega(b) \right\},
\]
where \(\Omega\) is some penalty function and \(\lambda>0\) is a regularization parameter.

@@

The term Bayesian inference usually refers to a wide class of methods that to some extent employ Bayes' theorem to update the initial probability assigned to a hypothesis (or a parameter) when new information is available. It can be seen as a general tool for modeling situations or problems that involve uncertainty; some examples include Bayesian hierarchical modeling, Bayesian regression or even Bayesian neural networks \citep[see e.g.][]{murphy2012machine, bishop2006pattern}. Specifically, in this work we will be interested in performing parameter estimation in a Bayesian framework, so that pre-existing information and beliefs about the parameters can be incorporated into the model.

@@

In this section we present a brief exposition of some basic concepts regarding reproducing kernel Hilbert spaces. Although there are quite a few ways of introducing these spaces, we adopt a probabilistic point of view that will set the stage for the development of the subsequent theory. For a more detailed account, a good reference is the book by~\citet{berlinet2004reproducing}. Although the following ideas can be extended to complex functions, we restrict ourselves to real-valued functions for the sake of simplicity. We start by defining the concept of kernel functions, which are the foundation of these spaces.

@@

Suppose \(X=X(t)\) is a \(L^2\)-stochastic process with trajectories in \(L^2[0, 1]\), and for simplicity assume \(\E[X(t)]=0\) for all \(t\in[0,1]\). Let us denote by \(K(t, s)= \E[X(t)X(s)]\) the covariance function of the process \(X\). To construct the RKHS \(\Hcal(K)\) associated with the covariance function, we start by defining the functional vector space \(\Hcal_0(K)\) of all finite linear combinations of evaluations of \(K\), that is,
\begin{equation}\label{eq:h0}
\Hcal_0(K) = \left\{ f: \ f(\cdot) = \sum_{i=1}^p a_i K(t_i, \cdot), \ p \in \N, \ a_i \in \R, \ t_i \in [0, 1] \right\}.
\end{equation}
Note that, as subsets, \(\Hcal_0(K)\subset L^2[0,1]\). However, this new space can be endowed with an inner product different from the one induced by \(L^2[0,1]\), namely
\[
\dotprod{f}{g}_K = \sum_{i, j} a_i b_j K(t_i, s_j),
\]
for \(f(\cdot)=\sum_i a_i K(t_i, \cdot) \) and \(g(\cdot)=\sum_j b_j K(s_j, \cdot)\). We show below that it is well defined, but before we note that functions in this space satisfy the so-called \textit{reproducing property}: \(f(t) = \dotprod{K(t, \cdot)}{f}_K\) for all \(t \in [0, 1]\). In particular, \(K(t, s)=\dotprod{K(t, \cdot)}{K(s, \cdot)}_K\), which can be understood as saying that the kernel \textit{reproduces} itself, hence the term ``reproducing kernel''.

@@

\begin{example}[Non-RKHS Hilbert space]
  As we pointed out before, the space of square integrable functions \(L^2(\R)\) \textit{is not} a RKHS. There are many ways of seeing this; for example, in this space norm convergence does not imply pointwise convergence. Another possibility is to observe that the would-be reproducing kernel must be the Dirac delta function:
  \[
    f(s) = \int_{-\infty}^\infty f(t)\delta(t, s)\, dt \quad \forall s \in \R.
  \]
  However, \(\delta(t, \cdot) \notin L^2(\R)\).
\end{example}

@@

Although not the main concern of this work, the theory of reproducing kernels and the associated kernel methods find applications in many areas of machine learning. For example, the well-known \textit{kernel trick} can be seen as a specific application of the reproducing property, where a feature map \(\Phi(x) \in \Hcal\) is applied to the objects in the space of interest, transforming them into richer elements in a RKHS. Then, many computations can be efficiently carried out in an implicit manner by means of the corresponding kernel
\[
K(x, y)=\dotprod{\Phi(x)}{\Phi(y)}_{\Hcal}.
\]
The fact that one does not need to explicitly compute \(\Phi(x)\) is especially relevant for kernels with a simple expression but a possibly complex feature space (e.g. infinite-dimensional), such as the Gaussian kernel \(K(x, y)=\exp(-\gamma\|x-y\|^2)\).

@@

\begin{theorem}[Generalized representer theorem]
  Consider a RKHS of functions \(\Hcal(K)\) defined over a non-empty set \(\X\), a training set \(\{(x_i, y_i)\}_{i=1}^n \subset (\X \times \R)^n\), an arbitrary error function \(R: (\X  \times \R^2)^n \to \R\cup\{\infty\}\), and a strictly increasing real-valued function \(\Omega:[0,\infty)\to \R\). Then, any minimizer \(f^\ast\) of the regularized empirical risk
  \[
    R(\{(x_i, y_i, f(x_i))\}_{i=1}^n) + \Omega(\|f\|_K)
  \]
  is of the form \(f^\ast(\cdot) = \sum_{i=1}^n a_i K(x_i, \cdot)\), where \(a_i\in\R\) for all \(i=1,\dots,n\).
\end{theorem}

@@

The Metropolis-Hastings algorithm \citep{metropolis1953equation} is arguably the best known algorithm in this context. As a matter of fact, it is actually more of a general framework from which many other methods are derived. In its original formulation, it relies on a symmetric \textit{proposal} distribution \(g(x'|x_t)\) that generates a proposal for the next state given the current one (e.g. a Gaussian distribution centered on \(x_t\)). Approximate samples from the intended distribution can then be generated through the following iterative acceptance-based procedure:
\begin{enumerate}[1.]
  \item Generate a proposal \(x' \sim g(x'|x_t)\).
  \item Accept the proposal with probability \(\alpha=\min\{1, f(x')/f(x_t)\}\). If accepted, set \(x_{t+1}=x'\); otherwise set \(x_{t+1}=x_t\).
\end{enumerate}

@@

Figure~\ref{fig:mh} shows a schematic representation of the process. Note that \(f(x')/f(x_t)=\pi(x')/\pi(x_t)\) is a measure of how more likely is \(x'\) to be a sample from \(\pi\) than \(x_t\). This specific acceptance rate is also chosen so that the detailed balance equations of the chain are satisfied, i.e., \(\pi(x)P(x'|x)=\pi(x')P(x|x')\), where \(P(x|y)\) represents the transition probability from state \(x\) to state \(y\). These conditions essentially guarantee that \(\pi\) is the stationary distribution of the chain; see \citet{robert1999monte} for additional details. An immediate generalization consists on choosing a general (possibly non-symmetric) jump distribution \(g\), in which case the acceptance probability is rewritten as
\[
\alpha = \min\left\{1, \frac{f(x')g(x_t \mid x')}{f(x_t)g(x'\mid x_t)}\right\}.
\]

Although this algorithm performs well on a wide range of problems, it can be difficult to find a suitable jump distribution when the number of dimensions is high. If we know the conditional distribution of each variable on the rest of them, an alternative approach to sampling is \textit{Gibb's algorithm} \citep{geman1984stochastic}, which is a particular case of Metropolis-Hastings that considers separate samples for each dimension. Specifically, in each step we get unidimensional samples \textit{in turn} (e.g. via rejection sampling) of each variable conditional on the rest of them, using always the more up-to-date values available:

### 2022-09-04 chapters/bayesian-functional-regression.tex
Finish first complete revision

  \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) that acts as a summary of the marginal posterior distributions, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\), and then predict the responses in the usual way following model~\eqref{eq:rkhs-model-linear}, i.e.:
  \[
    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
  \]
  Note that in this case the variance \(\sigma^2\) is treated as a nuisance parameter. Although it contributes to measure the uncertainty in the approximations, its estimates are discarded in the final prediction.

@@

With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
\[
\hat Y_i = \I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
\]
where \(\I\) is the indicator function (\(\I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise). The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.

On the other hand, the prediction method that takes into account the entire posterior approximation (i.e. the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples. Hence, there are primarily two possible outcomes.

>>>>

  \paragraph{Summarize-then-predict.} If we consider a point-estimate statistic \(T\) that acts as a summary of the marginal posterior distributions, we can get the corresponding estimates \(\hat{\theta}=(\hat b, \hat \tau, \hat{\alpha}_0, \hat{\sigma}^2) = T\{\theta^{(m)*}\} \equiv (T\{b^{(m)*}\}, T\{\tau^{(m)*}\}, T\{\alpha_0^{(m)*}\}, T\{(\sigma^2)^{(m)*}\})\), and then predict the responses in the usual way following model~\eqref{eq:rkhs-model-linear}, i.e.:
  \begin{equation}\label{eq:summarize-predict-linear}
    \hat Y_i = \hat{\alpha}_0 + \sum_{j=1}^p \hat{\beta}_j X_i(\hat{t}_j), \quad i=1,\dots, n'.
  \end{equation}
  Note that in this case the variance \(\sigma^2\) is treated as a nuisance parameter. Although it contributes to measure the uncertainty in the approximations, its estimates are discarded in the final prediction.

@@

With this last strategy in mind, the summarize-then-predict approach on the approximate posterior distribution is analogous to the linear regression case:
\begin{equation}\label{eq:summarize-predict-logistic}
\hat Y_i = \I \left( \left[\displaystyle 1 + \exp\left\{-\hat\alpha_0 - \sum_{j=1}^p \hat\beta_j X_i(\hat t_j)\right\}\right]^{-1} \geq \gamma \right), \quad i=1,\dots,n',
\end{equation}
where \(\I\) is the indicator function (\(\I(P)\) is \(1\) if \(P\) is true and \(0\) otherwise). The hat estimates are obtained once again through a summary statistic \(T\) of the corresponding marginal posterior distributions.

On the other hand, the prediction method that takes into account the entire posterior approximation (that is, the predict-then-summarize approach) is somewhat different now, since there is the question of which response (the Bernoulli variables in~\eqref{eq:rkhs-model-logistic-2} or the raw probabilities in~\eqref{eq:rkhs-model-logistic-2-parameter}) to consider when averaging the posterior samples. Hence, there are primarily two possible outcomes.

### 2022-09-04 chapters/conclusions.tex
Finish first complete revision

In this work we have addressed some inference and prediction issues involved in the task of learning from functional data. In particular, we have tackled functional linear and logistic regression problems, and we have presented an alternative to the \(L^2\)-models predominant in these scenarios. Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations from a functional perspective that are conceptually simpler and facilitate the interpretation and subsequent analysis, an attribute that practitioners surely find especially compelling. The underlying theoretical aspects are nevertheless far from trivial, so we have also explored some of the related mathematical background in this text.

With the aid of Bayesian methods, we have introduced a computationally feasible approach to parameter estimation within the proposed RKHS models. The flexibility is improved through the use of prior distributions that can introduce pre-existing information into the model. In addition, the fact that we have a complete posterior distribution at our disposal enables us to derive different prediction and variable selection methods, or even perform other types of inference such as the construction of credible intervals. Moreover, the use of MCMC methods to approximate the posterior distribution results in an implementation that is simpler than most existing alternatives. Although this simplicity comes at the expense of increased computational requirements, this computational cost is still reasonable and the model can be used in practice.

>>>>

In this work we have addressed some inference and prediction issues involved in the task of learning from functional data. In particular, we have tackled functional linear and logistic regression problems, and we have presented an alternative to the \(L^2\)-models predominant in these scenarios. Our proposed models rely heavily on the theory of RKHS's, offering finite-dimensional approximations from a functional perspective. This leads to a conceptually simpler approach that facilitates the interpretation and subsequent analysis, which we believe is a compelling feature for most practitioners. The underlying theoretical aspects are nevertheless far from trivial, so we have also explored some of the related mathematical background in this text.

With the aid of Bayesian methods, we have introduced a computationally feasible approach to parameter estimation within the proposed RKHS models. The flexibility is improved through the use of prior distributions that can introduce pre-existing information into the model. In addition, the fact that we have a complete posterior distribution at our disposal enables us to derive different prediction and variable selection methods, or even perform other types of inference such as the construction of credible intervals. Moreover, the use of MCMC methods to approximate the posterior distribution results in an implementation that is simpler than most existing alternatives. Although this simplicity comes at the expense of increased computational requirements, this computational cost is still reasonable and the model can be used in practice. The execution time only starts to be of some relevance when performing extensive simulations and comparisons, but it is not a serious burden for the use of our methods in practical cases.

### 2022-09-04 chapters/experiments.tex
Finish first complete revision

First and foremost, we consider \(T=\{\text{mean, median, mode}\}\) for the summarize-then-predict approach to prediction. In this way in linear regression we have 4 prediction methods (one for each statistic and one for the predict-then-summarize approach, which will we call ``\textit{posterior\_mean}'' for short) and 3 variable selection methods (one for each statistic), while in logistic regression there are similarly 5 prediction methods and 3 variable selection procedures. In each case, after variable selection is performed, we use a penalized multiple linear/logistic regression method to generate the corresponding predictions. All in all, we are looking at 7 (8 in the case of logistic regression) prediction methods, and although all of them are derived from a single MCMC run, we will treat them as separate for the purposes of the experimentation. In what follows we relax the notation and use the expressions ``model'', ``prediction method'' and ``algorithm'' almost interchangeably when referring to these 7 (or 8) cases.

@@

We consider a set of functional regressors common to linear and logistic regression problems. They are 4 Gaussian processes (GP's), each with a different kernel function:
\begin{enumerate}
  \item BM.\hspace{0.3em} A Brownian motion \(K_1(t,s)=\min\{t,s\}\).
  \item fBM.\hspace{0.3em} A fractional Brownian motion \(K_2(t,s)=1/2(s^{2H} + t^{2H} - |t-s|^{2H})\), with Hurst parameter \(H=0.8\).
  \item O-U.\hspace{0.3em} An Ornstein-Uhlenbeck process \(K_3(t,s)=e^{-|t-s|}\).
  \item Gaussian.\hspace{0.3em} A Gaussian kernel \(K_4(t,s)=e^{-(t-s)^2/2\nu^2}\), with \(\nu=0.2\).
\end{enumerate}
Also, when applicable, we fix a variance \(\sigma^2=0.5\) for the error terms \(\epsilon\).

@@

The reference algorithms used were all fast methods that can complete their whole CV loop in just a few seconds, so any comparison with them in this regard is uninteresting. However, this outcome was to be expected, since in general the MCMC methods are known to have a high computational cost in exchange for the simplicity in modeling they provide. An individual MCMC run is not that expensive time-wise, but even for a modest 5-fold cross-validation loop the effects add up quickly to a considerable amount of time. In Figure~\ref{fig:dependence_time_n_train} we show the evolution of the execution time versus the number of training samples in a generic case, which not surprisingly has a more or less linear behavior.

>>>>

First and foremost, we consider \(T=\{\text{mean, median, mode}\}\) for our summarize-then-predict approach to prediction (see~\eqref{eq:summarize-predict-linear} and~\eqref{eq:summarize-predict-logistic}). As a result, in linear regression we have 4 prediction methods (one for each statistic and one for the predict-then-summarize approach) and 3 variable selection methods (one for each statistic), while in logistic regression there are similarly 5 prediction methods and 3 variable selection procedures. In each case, after variable selection is performed, we use a \(l^2\)-penalized multiple linear/logistic regression method to generate the corresponding predictions. All in all, we are looking at 7 (8 in the case of logistic regression) prediction methods, and although all of them are derived from a single MCMC run, we will treat them as separate for the purposes of the experimentation. In what follows we relax the notation and use the expressions ``model'', ``prediction method'' and ``algorithm'' almost interchangeably when referring to these 7 (or 8) cases.

@@

We consider a set of functional regressors common to linear and logistic regression problems. They are 4 Gaussian processes (GP's), each with a different covariance function:
\begin{enumerate}
  \item BM.\hspace{0.3em} A Brownian motion \(K_1(t,s)=\min\{t,s\}\).
  \item fBM.\hspace{0.3em} A fractional Brownian motion \(K_2(t,s)=1/2(s^{2H} + t^{2H} - |t-s|^{2H})\), with Hurst parameter \(H=0.8\).
  \item O-U.\hspace{0.3em} An Ornstein-Uhlenbeck process \(K_3(t,s)=e^{-|t-s|}\).
  \item Gaussian.\hspace{0.3em} A Gaussian process with Gaussian kernel \(K_4(t,s)=e^{-(t-s)^2/2\nu^2}\), where \(\nu=0.2\).
\end{enumerate}
Also, when applicable, we fix a variance \(\sigma^2=0.5\) for the error terms \(\epsilon\).

@@

The reference algorithms used were all fast methods that can complete their whole CV loop in just a few seconds, so any comparison with them in this regard is uninteresting. However, this outcome was to be expected, since in general the MCMC methods are known to have a high computational cost in exchange for the simplicity in modeling they provide. An individual MCMC run is not that expensive time-wise, but even for a modest 5-fold cross-validation loop the effects add up quickly to a considerable amount of time. In Figure~\ref{fig:dependence_time_n_train} we show the evolution of the execution time versus the number of training samples in a generic case, which not surprisingly has a more or less linear behavior. This, however, is not an impediment for using our methods in practice.
